Double Standards in Judging Collective Action\thanks{}

### 2022-09-20 Documents/manuscript.Rmd
submit manuscript

---
title: "Double Standards in Judging Collective Action"
header: "DOUBLE STANDARDS"
keywords: protest, collective action, double standards, social identity theory, system justification theory, Black Lives Matter
abstract: "Collective action is a powerful force driving social change but often sparks contention about what actions are acceptable means to effect social change. In five studies (total *N* = 2,979), we investigated double standards in judging collective action---that is, whether observers will judge the same protest actions as more acceptable if the protesters are ingroup members and their cause aligns with the ingroup's interests (*identity-based double standards*) or if the protesters' cause aligns with the observers' ideological positions (*ideology-based double standards*). In two studies, we developed an instrument of 25 controversial protest actions, based in item response theory, to measure where people draw the line between acceptable and unacceptable forms of collective action. In three preregistered experiments, we used this instrument to test for double standards in judging collective action for workers' rights in the United Kingdom and for or against defunding the police and for or against restricting abortion in the United States. We found evidence for ideology-based, but not for identity-based, double standards: Participants judged the same protest actions to be more acceptable when the protesters' cause aligned with their own ideological position---but showed no consistent ingroup bias when judging collective action by ingroup and outgroup members. Our findings have theoretical and practical implications for understanding the often divided response to prominent social movements."
bibliography: references.bib
csl: apa.csl
output:
  pdf_document:
    latex_engine: pdflatex
    keep_tex: yes
    template: apa-template.tex

@@

@@

In Experiment 2, we investigated potential double standards in judging collective action for or against defunding the police in the United States. That is, we tested whether Black and White Americans applied different standards when judging collective action taken by either Black or White protesters to protest either for or against police divestment as a possible solution to racist police violence. We conducted this study in January 2021, a time when, after months of unprecedented protests for racial justice, most Americans could be expected to be aware of, and have formed an opinion on, the Black Lives Matter movement. We focused on defunding the police as this position remained controversial among both Black and White Americans even as they broadly agreed on the need for police reform.^[In December 2020, only 34% of Black Americans and 18% of White Americans supported defunding the police while majorities of Black and White Americans supported police reforms such as eliminating qualified immunity (50% and 57%) or banning chokeholds [79% and 61%, @yougov_police_2020].]

>>>>

---
title: "Double Standards in Judging Collective Action"
header: "DOUBLE STANDARDS"
keywords: collective action, double standards, social identity theory, system justification theory, Black Lives Matter
abstract: "Collective action is a powerful force driving social change but often sparks contention about what actions are acceptable means to effect social change. In five studies (total *N* = 2,979), we investigated double standards in judging collective action---that is, whether observers will judge the same protest actions as more acceptable if the protesters are ingroup members and their cause aligns with the ingroup's interests (*identity-based double standards*) or if the protesters' cause aligns with the observers' ideological positions (*ideology-based double standards*). In two studies, we developed an instrument of 25 controversial protest actions, based in item response theory, to measure where people draw the line between acceptable and unacceptable forms of collective action. In three preregistered experiments, we used this instrument to test for double standards in judging collective action for workers' rights in the United Kingdom and for or against defunding the police and for or against restricting abortion in the United States. We found evidence for ideology-based, but not for identity-based, double standards: Participants judged the same protest actions to be more acceptable when the protesters' cause aligned with their own ideological position---but showed no consistent ingroup bias when judging collective action by ingroup and outgroup members. Our findings have theoretical and practical implications for understanding the often divided response to prominent social movements."
bibliography: references.bib
csl: apa.csl
output:
  pdf_document:
    latex_engine: pdflatex
    keep_tex: yes
    template: apa-template.tex

@@

### Transparency and Openness

@@

In Experiment 2, we investigated potential double standards in judging collective action for or against defunding the police in the United States. That is, we tested whether Black and White Americans applied different standards when judging collective action taken by either Black or White protesters to protest either for or against police divestment as a possible solution to racist police violence. We conducted this study in January 2021, a time when, after months of unprecedented protests for racial justice, most Americans could be expected to be aware of, and have formed an opinion on, the Black Lives Matter movement [@leach_social_2017; @leach_understanding_2022]. We focused on defunding the police as this position remained controversial among both Black and White Americans even as they broadly agreed on the need for police reform.^[In December 2020, only 34% of Black Americans and 18% of White Americans supported defunding the police while majorities of Black and White Americans supported police reforms such as eliminating qualified immunity (50% and 57%) or banning chokeholds [79% and 61%, @yougov_police_2020].]

### 2022-09-20 Documents/manuscript.tex
submit manuscript

@@

In Experiment 2, we investigated potential double standards in judging
collective action for or against defunding the police in the United
States. That is, we tested whether Black and White Americans applied
different standards when judging collective action taken by either Black
or White protesters to protest either for or against police divestment
as a possible solution to racist police violence. We conducted this
study in January 2021, a time when, after months of unprecedented
protests for racial justice, most Americans could be expected to be
aware of, and have formed an opinion on, the Black Lives Matter
movement. We focused on defunding the police as this position remained
controversial among both Black and White Americans even as they broadly
agreed on the need for police reform.\footnote{In December 2020, only
  34\% of Black Americans and 18\% of White Americans supported
  defunding the police while majorities of Black and White Americans
  supported police reforms such as eliminating qualified immunity (50\%
  and 57\%) or banning chokeholds (79\% and 61\%, YouGov, 2020).}

@@

@@

@@

>>>>

\hypertarget{transparency-and-openness}{%
\subsubsection{Transparency and
Openness}\label{transparency-and-openness}}

@@

In Experiment 2, we investigated potential double standards in judging
collective action for or against defunding the police in the United
States. That is, we tested whether Black and White Americans applied
different standards when judging collective action taken by either Black
or White protesters to protest either for or against police divestment
as a possible solution to racist police violence. We conducted this
study in January 2021, a time when, after months of unprecedented
protests for racial justice, most Americans could be expected to be
aware of, and have formed an opinion on, the Black Lives Matter movement
(Leach \& Allen, 2017; Leach \& Teixeira, 2022). We focused on defunding
the police as this position remained controversial among both Black and
White Americans even as they broadly agreed on the need for police
reform.\footnote{In December 2020, only 34\% of Black Americans and 18\%
  of White Americans supported defunding the police while majorities of
  Black and White Americans supported police reforms such as eliminating
  qualified immunity (50\% and 57\%) or banning chokeholds (79\% and
  61\%, YouGov, 2020).}

@@

\hypertarget{transparency-and-openness-1}{%
\subsubsection{Transparency and
Openness}\label{transparency-and-openness-1}}

@@

\hypertarget{transparency-and-openness-2}{%
\subsubsection{Transparency and
Openness}\label{transparency-and-openness-2}}

@@

\leavevmode\vadjust pre{\hypertarget{ref-leach_social_2017}{}}%
Leach, C. W., \& Allen, A. M. (2017). The social psychology of the
{Black Lives Matter} meme and movement. \emph{Current Directions in
Psychological Science}, \emph{26}(6), 543--547.
\url{https://doi.org/10.1177/0963721417719319}

\leavevmode\vadjust pre{\hypertarget{ref-leach_understanding_2022}{}}%
Leach, C. W., \& Teixeira, C. P. (2022). Understanding sentiment toward
{``{Black Lives Matter}.''} \emph{Social Issues and Policy Review},
\emph{16}(1), 3--32. \url{https://doi.org/10.1111/sipr.12084}
